Binocular integration of chromatic and luminance signals

### 2024-07-16 manuscript.tex
Testing kbl instead of kable for table production
Revised in line with reviewer comments

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\KOMAoption{captions}{tableheading}
\usepackage{float} \floatplacement{figure}{H} \newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{A\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{A\arabic{figure}}}
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Binocular integration of chromatic and luminance signals},
  pdfauthor={Daniel H. Baker\^{}\{1,2,*\}; Kirralise J. Hansford\^{}1; Federico G. Segala\^{}1; Annie Y. Morsi\^{}1; Rowan J. Huxley\^{}\{1,3\}; Joel T. Martin\^{}\{1,4\}; Maya Rockman\^{}1; Alex R. Wade\^{}\{1,2\}},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

@@

\begin{document}
\maketitle
\(^1\)Department of Psychology, University of York, UK\\
\(^2\)York Biomedical Research Institute, University of York, UK\\
\(^3\)School of Psychology, University of Nottingham, UK\\
\(^4\)School of Philosophy, Psychology and Language Sciences, University
of Edinburgh, UK\\
\(^*\)Corresponding author, email: daniel.baker@york.ac.uk

\hypertarget{abstract}{%
\section{Abstract}\label{abstract}}

Much progress has been made in understanding how the brain combines
signals from the two eyes. However, most of this work has involved
achromatic (black and white) stimuli, and it is not clear if the same
processes apply in colour-sensitive pathways. In our first experiment,
we measured contrast discrimination (`dipper') functions for four key
ocular configurations (monocular, binocular, half-binocular and
dichoptic), for achromatic, isoluminant red/green and isoluminant
blue/yellow sine-wave grating stimuli. We find a similar pattern of
results across stimuli, implying equivalently strong interocular
suppression within each pathway. Our second experiment measured
dichoptic masking within and between pathways using the method of
constant stimuli. Masking was strongest within-pathway, and weakest
between blue/yellow and achromatic mechanisms. Finally, we repeated the
dipper experiment using temporal luminance modulations, which produced
slightly weaker interocular suppression than for spatially modulated
stimuli. We interpret our results in the context of a contemporary
two-stage model of binocular contrast gain control, implemented here
using a hierarchical Bayesian framework. Posterior distributions of the
weight of interocular suppression overlapped with a value of 1 for all
dipper data sets, and the model captured well the pattern of thresholds
from all three experiments.

@@

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

The process by which the brain combines independent inputs is of
fundamental importance for understanding sensory perception. Binocular
vision is a useful test-case for determining the general principles
involved in neural signal combination, as our brains typically combine
the inputs from the left and right eyes to provide binocular single
vision. In recent years our understanding has been facilitated by the
development of binocular gain control models that provide a framework to
interpret empirical data from multiple experimental paradigms and
techniques, including psychophysics (Meese et al., 2006), EEG (Baker \&
Wade, 2017), fMRI (Moradi \& Heeger, 2009) and pupillometry (Segala et
al., 2023). However, most of this work has used achromatic (black and
white) stimuli; we know comparatively little about how chromatic signals
are combined binocularly, or about how signals in different ocular and
chromatic channels interact. In this study we use psychophysical
detection and discrimination paradigms to explore binocular interactions
within and between the chromatic and achromatic pathways.

A useful framework for understanding binocular signal processing is the
two-stage gain control model of binocular combination introduced by
Meese et al. (2006). This model features interocular suppression between
monocular channels, followed by binocular summation. The model accounts
well for the pattern of contrast discrimination (`dipper') functions for
four distinct ocular configurations (see also Georgeson et al., 2016),
illustrated in Figure~\ref{fig-exampledips}. In the monocular condition,
participants must discriminate between stimuli of two contrasts (a
`pedestal', and a `pedestal plus target') that are both presented to one
eye, whilst the other eye views mean luminance. Threshold is defined as
the minimum target contrast required to make this judgement with 75\%
accuracy. The binocular condition is the same, except that the stimuli
are shown to both eyes. In the half-binocular condition the pedestal is
shown to both eyes, but the target increment is shown only to one eye.
Finally, the dichoptic condition involves presenting the pedestal to one
eye, and the target increment to the other eye.

@@

{\centering \includegraphics{Figures/exampledips.pdf}

@@

\end{figure}

The detailed pattern of thresholds across these four conditions is
complex, and for achromatic stimuli has several distinctive features
that have been replicated in multiple studies. At low pedestal
contrasts, the binocular condition yields lower thresholds than the
monocular condition; this result is attributed to physiological
binocular summation by neurons responsive to signals from both eyes
(Baker et al., 2018; Campbell \& Green, 1965). However, at high pedestal
contrasts the `handle' regions of the dipper functions for these
conditions converge: a consequence of interocular suppression
compensating for the increased excitation during binocular stimulation
(Legge, 1984; Maehara \& Goryo, 2005). The half-binocular condition
avoids confounding the number of eyes seeing the target with the number
of eyes seeing the pedestal (Meese et al., 2006). The pedestal is always
binocular in this condition, whereas the target increment is monocular,
and thresholds are consistently higher than in the binocular condition
across the full range of pedestal contrasts. This demonstrates that
binocular summation occurs across the full contrast range, when the
pedestal ocularity is appropriately controlled. Finally, the dichoptic
condition produces extremely strong masking of the target, such that
when the pedestal is visible, the target must equal or exceed its
contrast to be detectable (Baker \& Meese, 2007; Legge, 1979; Maehara \&
Goryo, 2005). The characteristic pattern of dipper functions (right
panel of Figure~\ref{fig-exampledips}) is well-described by the gain
control model of Meese et al. (2006) for achromatic stimuli.

At the output of the human retina, cone responses are split into three
distinct pathways. The sum of long- and medium-wavelength cone outputs
(L+M) transmits luminance information (see Sharpe et al., 2011), and is
likely responsible for the binocular combination effects previously
studied using achromatic stimuli (see above). The difference of long-
and medium-wavelength cone outputs (L-M) is responsive to chromatic
stimuli modulating along a red/green axis in colour space. Finally the
opponent (S-(L+M)) channel codes chromatic stimuli modulating along a
blue/yellow axis. At isoluminance, where (L+M) is held constant, this
channel is driven purely by the S-cone outputs.

There has not yet been a detailed investigation of binocular contrast
discrimination in either of these chromatic pathways, however there is
reason to believe they may differ from the achromatic pathway. At
detection threshold, binocular summation is greater for chromatic versus
achromatic stimuli (Simmons, 2005), implying a more linear initial stage
of processing. For cross-orientation masking, there are differences in
the magnitude of masking between chromatic (red/green) and achromatic
stimuli (Kim et al., 2013; Medina \& Mullen, 2009), as well as
differences in their temporal dynamics (Kim \& Mullen, 2015). There are
also interactions between chromatic and achromatic pathways both within
(Chen et al., 2000) and between (Kingdom \& Libenson, 2015; Mullen et
al., 2014) the eyes, yet these have not been fully explored for
arrangements where the target and mask have the same orientation.
Finally, the neurophysiological underpinnings of colour vision are
distinct from those of the achromatic system. In primary visual cortex
(V1), chromatic signals are processed in `blob' regions that are
revealed by cytochrome oxidase staining (Horton \& Hubel, 1981). The
blob regions appear to be largely monocular (Livingstone \& Hubel,
1984), suggesting that binocular combination for chromatic stimuli might
occur separately than for achromatic stimuli, and perhaps be subject to
different constraints.

A further property of `blob' regions is that they are less strongly
orientation-tuned than other regions of V1 (Horton \& Hubel, 1981;
Livingstone \& Hubel, 1984), and biased towards low spatial frequencies
(Edwards et al., 1995; Tootell et al., 1988; though see Economides et
al., 2011). Our recent work (Segala et al., 2023) has investigated
binocular combination for flickering discs of luminance, which are
DC-balanced across time (i.e.~the time-averaged luminance is equal to
the background). Steady-state EEG responses from early visual cortex and
psychophysical contrast matching data were both consistent with weak
interocular suppression when using this stimulus arrangement. This is
very different from the `ocularity invariance' (where binocular and
monocular stimuli appear equal) that is well-established when using
spatially modulated stimuli such as sine-wave gratings, and implies
strong interocular suppression (Baker \& Wade, 2017; Meese et al., 2006;
Moradi \& Heeger, 2009). There is additional psychophysical evidence
from matching studies that binocular combination can be close to linear
for luminance increments (Anstis \& Ho, 1998; Levelt, 1965),
particularly against a dark background (Baker et al., 2012). Both
structural and functional data therefore imply that binocular
combination may differ between spatial and temporal contrast.

The main aim of the present study was to characterise binocular signal
combination for chromatic stimuli, and for temporal modulations of
luminance. We also aimed to investigate interocular suppression between
chromatic and achromatic pathways. We therefore preregistered a series
of psychophysical experiments (see: \url{https://osf.io/3vdga/}). In
Experiment 1 we replicate the four key pedestal masking conditions of
Meese et al. (2006) described above for achromatic grating stimuli and
extend this to both red/green and blue/yellow isoluminant chromatic
stimuli. In Experiment 2 we explore dichoptic masking within and between
these stimuli. Experiment 3 repeats the achromatic condition from the
first experiment but using a temporally modulated disc rather than
sine-wave gratings. We take a Bayesian approach to data analysis and
modelling; by fitting a hierarchical version of the two-stage gain
control model (Meese et al., 2006) we compare posterior parameter
distributions to understand how model parameters such as the weight of
interocular suppression vary across visual pathways.

\hypertarget{materials-methods}{%
\section{Materials \& Methods}\label{materials-methods}}

\hypertarget{participants}{%
\subsection{Participants}\label{participants}}

All experiments were completed by the first author (DHB) and two
additional participants, who differed for each experiment. Participants
had no known abnormalities of binocular or colour vision. Written
informed consent was obtained before data collection began, and all
procedures were approved by the ethics committee of the Department of
Psychology at the University of York (ID number 2202).

\hypertarget{apparatus-stimuli}{%
\subsection{Apparatus \& stimuli}\label{apparatus-stimuli}}

In Experiments 1 and 2, the stimuli were horizontal sinusoidal gratings
with a spatial frequency of 1c/deg (see examples in
Figure~\ref{fig-examplestim}). The gratings were windowed by a raised
cosine envelope with a diameter of 3 degrees. Spatial phase, relative to
a central fixation cross, was randomised on each trial across the four
cardinal phases. In the achromatic conditions, the sine-wave modulated
all three monitor colour channels equally (red, green and blue). In the
L-M condition, we generated isoluminant stimuli for each participant
(see Procedures) designed to maximise contrast between L and M cones,
whilst keeping S cone activity constant. In the S-(L+M) condition, the
isoluminant stimuli maximised S-cone contrast. Stimuli were converted
from cone space to monitor RGB coordinates using the monitor spectral
readings and the Stockman-Sharpe 2 degree cone fundamentals (Stockman \&
Sharpe, 2000). The maximum displayable cone contrasts on our system were
0.1 for L-M, and 0.88 for S-(L+M). The stimuli in Experiment 3 were
temporal modulations of luminance applied to a disc made using the same
raised cosine envelope as described above, but with no further spatial
modulation. The stimuli counterphase flickered sinusoidally at 4Hz (see
examples in the lower portion of Figure~\ref{fig-examplestim}). In all
experiments, we displayed a binocular fusion lock, consisting of three
concentric rings of small square elements with random colour. A black
central fixation cross was also displayed throughout.

@@

{\centering \includegraphics[width=0.5\textwidth,height=\textheight]{Figures/examplestim.pdf}

@@

\end{figure}

@@

\hypertarget{procedure}{%
\subsection{Procedure}\label{procedure}}

@@

{\centering \includegraphics{Figures/isosettings.pdf}

@@

\caption{\label{fig-isofig}Isoluminance settings from all participants
in Experiments 1 and 2. Panel (a) shows red/green and panel (b) shows
blue/yellow settings that were subsequently used to generate stimuli in
the main experiments. Within each panel, solid lines show the mean
settings for each participant, and black curves show the range of
possible stimuli displayed during the adjustment task.}

\end{figure}

@@

\hypertarget{data-analysis-and-computational-modelling}{%
\subsection{Data analysis and computational
modelling}\label{data-analysis-and-computational-modelling}}

@@

\begin{equation}\protect\hypertarget{eq-s1l}{}{\mathrm{Stage1}_L = \frac{C_L^m}{S + C_L + \omega C_R},}\label{eq-s1l}\end{equation}

\begin{equation}\protect\hypertarget{eq-s1r}{}{\mathrm{Stage1}_R = \frac{C_R^m}{S + C_R + \omega C_L},}\label{eq-s1r}\end{equation}

\begin{equation}\protect\hypertarget{eq-binsum}{}{\mathrm{binsum} = \mathrm{Stage1}_L + \mathrm{Stage1}_R,}\label{eq-binsum}\end{equation}

\begin{equation}\protect\hypertarget{eq-s2}{}{\mathrm{Stage2} = \frac{\mathrm{binsum}^p}{Z + \mathrm{binsum}^q},}\label{eq-s2}\end{equation}

>>>>

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
% definitions for citeproc citations
\NewDocumentCommand\citeproctext{}{}
\NewDocumentCommand\citeproc{mm}{%
  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
\makeatletter
 % allow citations to break across lines
 \let\@cite@ofmt\@firstofone
 % avoid brackets around text for \cite:
 \def\@biblabel#1{}
 \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
 {\begin{list}{}{%
  \setlength{\itemindent}{0pt}
  \setlength{\leftmargin}{0pt}
  \setlength{\parsep}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
   \setlength{\leftmargin}{\cslhangindent}
   \setlength{\itemindent}{-1\cslhangindent}
  \fi
  % set entry spacing
  \setlength{\itemsep}{#2\baselineskip}}}
 {\end{list}}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\KOMAoption{captions}{tableheading}
\usepackage{lineno} \linenumbers \usepackage{float} \floatplacement{figure}{H} \newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{A\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{A\arabic{figure}}}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage{bookmark}

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Binocular integration of chromatic and luminance signals},
  pdfauthor={Daniel H. Baker\^{}\{1,2,*\}; Kirralise J. Hansford\^{}1; Federico G. Segala\^{}1; Annie Y. Morsi\^{}1; Rowan J. Huxley\^{}\{1,3\}; Joel T. Martin\^{}\{1,4\}; Maya Rockman\^{}1; Alex R. Wade\^{}\{1,2\}},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

@@

\begin{document}
\maketitle

\(^1\)Department of Psychology, University of York, UK\\
\(^2\)York Biomedical Research Institute, University of York, UK\\
\(^3\)School of Psychology, University of Nottingham, UK\\
\(^4\)School of Philosophy, Psychology and Language Sciences, University
of Edinburgh, UK\\
\(^*\)Corresponding author, email: daniel.baker@york.ac.uk

\section{Abstract}\label{abstract}

Much progress has been made in understanding how the brain combines
signals from the two eyes. However, most of this work has involved
achromatic (black and white) stimuli, and it is not clear if the same
processes apply in colour-sensitive pathways. In our first experiment,
we measured contrast discrimination (`dipper') functions for four key
ocular configurations (monocular, binocular, half-binocular and
dichoptic), for achromatic, isoluminant L-M and isoluminant S-(L+M)
sine-wave grating stimuli. We find a similar pattern of results across
stimuli, implying equivalently strong interocular suppression within
each pathway. Our second experiment measured dichoptic masking within
and between pathways using the method of constant stimuli. Masking was
strongest within-pathway, and weakest between S-(L+M) and achromatic
mechanisms. Finally, we repeated the dipper experiment using temporal
luminance modulations, which produced slightly weaker interocular
suppression than for spatially modulated stimuli. We interpret our
results in the context of a contemporary two-stage model of binocular
contrast gain control, implemented here using a hierarchical Bayesian
framework. Posterior distributions of the weight of interocular
suppression overlapped with a value of 1 for all dipper data sets, and
the model captured well the pattern of thresholds from all three
experiments.

@@

\section{Introduction}\label{introduction}

The process by which the brain combines independent inputs is of
fundamental importance for understanding sensory perception (Baker \&
Wade, 2017; Ernst \& Banks, 2002). Binocular vision is a useful
test-case for determining the general principles involved in neural
signal combination, as our brains typically combine the inputs from the
left and right eyes to provide binocular single vision (Read, 2021). In
recent years our understanding has been facilitated by the development
of binocular gain control models that provide a framework to interpret
empirical data from multiple experimental paradigms and techniques,
including psychophysics (Meese et al., 2006), EEG (Baker \& Wade, 2017),
fMRI (Moradi \& Heeger, 2009) and pupillometry (Segala et al., 2023).
However, most of this work has used achromatic (black and white)
stimuli; we know comparatively little about how chromatic signals are
combined binocularly, or about how signals in different ocular and
chromatic channels interact. In this study we use psychophysical
detection and discrimination paradigms to explore binocular interactions
within and between the chromatic and achromatic pathways.

A useful framework for understanding binocular signal processing is the
two-stage gain control model of binocular combination introduced by
Meese et al. (2006). This model features interocular suppression between
monocular channels, followed by binocular summation. The model accounts
well for the pattern of contrast discrimination (`dipper') functions for
four distinct ocular configurations (see also Georgeson et al., 2016),
illustrated in Figure~\ref{fig-exampledips}. In the monocular condition,
participants must discriminate between stimuli of two contrasts (a
`pedestal', and a `pedestal plus target') that are both presented to one
eye, whilst the other eye views mean luminance. Threshold is defined as
the minimum target contrast required to make this judgement with 75\%
accuracy; this reduces at pedestal contrasts around threshold
(facilitation), and increases at high pedestal contrasts (masking). The
binocular condition is the same, except that the stimuli are shown to
both eyes. In the half-binocular condition the pedestal is shown to both
eyes, but the target increment is shown only to one eye. Finally, the
dichoptic condition involves presenting the pedestal to one eye, and the
target increment to the other eye.

@@

\centering{

\includegraphics{Figures/exampledips.pdf}

@@

\end{figure}%

The detailed pattern of thresholds across these four conditions is
complex, and for achromatic stimuli has several distinctive features
that have been replicated in multiple studies. At low pedestal
contrasts, the binocular condition yields lower thresholds than the
monocular condition; this result is attributed to physiological
binocular summation by neurons responsive to signals from both eyes
(Baker et al., 2018; Campbell \& Green, 1965). However, at high pedestal
contrasts the `handle' regions of the dipper functions for these
conditions converge: a consequence of interocular suppression
compensating for the increased excitation during binocular stimulation
(Legge, 1984; Maehara \& Goryo, 2005). The half-binocular condition
avoids confounding the number of eyes seeing the target with the number
of eyes seeing the pedestal (Meese et al., 2006). The pedestal is always
binocular in this condition, whereas the target increment is monocular,
and thresholds are consistently higher than in the binocular condition
across the full range of pedestal contrasts. This demonstrates that
binocular summation occurs across the entire contrast range, when the
pedestal ocularity is appropriately controlled. Finally, the dichoptic
condition produces extremely strong masking of the target, such that
when the pedestal is visible, the target must equal or exceed its
contrast to be detectable (Baker \& Meese, 2007; Legge, 1979; Maehara \&
Goryo, 2005). The characteristic pattern of dipper functions (right
panel of Figure~\ref{fig-exampledips}) is well-described by the gain
control model of Meese et al. (2006) for achromatic stimuli.

At the output of the human retina, cone responses are split into three
distinct pathways. The sum of long- and medium-wavelength cone outputs
(L+M) transmits luminance information (see Sharpe et al., 2011), and is
likely responsible for the binocular combination effects previously
studied using achromatic stimuli (see above). The difference of long-
and medium-wavelength cone outputs (L-M) is responsive to chromatic
stimuli modulating along a red/green axis in colour space. Finally the
opponent (S-(L+M)) channel codes chromatic stimuli modulating along a
blue/yellow axis. At isoluminance, where (L+M) is held constant, this
channel is driven purely by the S-cone outputs. Within each chromatic
pathway, contrast discrimination functions are similar to those for
achromatic vision, showing evidence of facilitation at low pedestal
contrasts and masking at high pedestal contrasts (Chen et al., 2000;
Switkes et al., 1988). There are also interactions between luminance and
colour mechanisms, which are able to mask each other (Chen et al., 2000;
Mullen \& Losada, 1994; Switkes et al., 1988), and under some
circumstances also show cross-pathway facilitation (Cole et al., 1990;
Mullen \& Losada, 1994; Shooner \& Mullen, 2020; Switkes et al., 1988).

There has not yet been a detailed investigation of binocular contrast
interactions in either chromatic pathway, however there are reasons to
believe they may differ from the achromatic pathway. At detection
threshold, binocular summation is greater for chromatic versus
achromatic stimuli (Simmons, 2005), implying a more linear initial stage
of processing. For cross-orientation masking, there are differences in
the magnitude of masking between chromatic (L-M) and achromatic stimuli
(Kim et al., 2013; Medina \& Mullen, 2009), as well as differences in
their temporal dynamics (Kim \& Mullen, 2015). There are also binocular
interactions between chromatic and achromatic pathways (Kingdom \&
Libenson, 2015; Mullen et al., 2014), yet these have not been fully
explored for arrangements where the target and mask have the same
orientation. Finally, the neurophysiological underpinnings of colour
vision may be distinct from those of the achromatic system. The
classical view is that in primary visual cortex (V1), chromatic signals
are processed in `blob' regions that are revealed by cytochrome oxidase
staining (Horton \& Hubel, 1981), and are largely monocular (Livingstone
\& Hubel, 1984). However subsequent work has revealed a more complex
arrangement, making such a clear-cut phyisological distinction unlikely
(for a review, see Shapley \& Hawken, 2011). Nevertheless, the
physiological segregation between pathways at the output of the retina
could still lead to differences in the way signals are combined across
the eyes.

The spatiotemporal profile of stimuli also appears to impact their
binocular combination. For example, matching studies indicate that
binocular combination can be close to linear for luminance increments
(Anstis \& Ho, 1998; Levelt, 1965), particularly against a dark
background (Baker et al., 2012). This is quite different from the
`ocularity invariance' (where binocular and monocular stimuli appear
equal) that is well-established when using spatially modulated stimuli
such as sine-wave gratings, and implies strong interocular suppression
(Baker \& Wade, 2017; Meese et al., 2006; Moradi \& Heeger, 2009). Our
recent work (Segala et al., 2023) has investigated binocular combination
for flickering discs of luminance, which are DC-balanced across time
(i.e.~the time-averaged luminance is equal to the background).
Steady-state EEG responses from early visual cortex and psychophysical
contrast matching data were both consistent with weak interocular
suppression when using this stimulus arrangement, but it is currently
unclear whether this has implications for contrast discrimination
performance.

The main aim of the present study was to characterise binocular signal
combination for chromatic stimuli, and for temporal modulations of
luminance. We also aimed to investigate interocular suppression between
chromatic and achromatic pathways. We therefore preregistered a series
of psychophysical experiments (see: \url{https://osf.io/3vdga/}). In
Experiment 1 we replicate the four key pedestal masking conditions of
Meese et al. (2006) described above for achromatic grating stimuli and
extend this to both L-M and S-(L+M) isoluminant chromatic stimuli. In
Experiment 2 we explore dichoptic masking within and between these
stimuli. Experiment 3 repeats the achromatic condition from the first
experiment but using a temporally modulated disc rather than sine-wave
gratings. We take a Bayesian approach to data analysis and modelling; by
fitting a hierarchical version of the two-stage gain control model
(Meese et al., 2006) we compare posterior parameter distributions to
understand how model parameters such as the weight of interocular
suppression vary across visual pathways.

\section{Materials \& Methods}\label{materials-methods}

\subsection{Participants}\label{participants}

All experiments were completed by the first author (DHB) and two
additional participants, who differed for each experiment. Participants
had no known abnormalities of binocular or colour vision. Written
informed consent was obtained before data collection began, and all
procedures were approved by the ethics committee of the Department of
Psychology at the University of York (ID number 2202). The study
conformed to the tenets of the Declaration of Helsinki.

\subsection{Apparatus \& stimuli}\label{apparatus-stimuli}

In Experiments 1 and 2, the stimuli were horizontal sinusoidal gratings
with a spatial frequency of 1c/deg (see examples in
Figure~\ref{fig-examplestim}). The gratings were windowed by a raised
cosine envelope with a diameter of 3 degrees. Spatial phase, relative to
a central fixation cross, was randomised on each trial across the four
cardinal phases. In the achromatic conditions, the sine-wave modulated
all three monitor colour channels equally (red, green and blue). In the
L-M condition, we generated isoluminant stimuli for each participant
(see Procedures) designed to maximise contrast between L and M cones,
whilst keeping S cone activity constant (more accurately, this is a
\(\Delta L/L - \Delta M/M\) condition, see Cole et al., 1993). In the
S-(L+M) condition (or more accurately the
\(\Delta S/S - (\Delta L/L + \Delta M/M)\) condition), the isoluminant
stimuli maximised S-cone contrast. Stimuli were converted from cone
space to monitor RGB coordinates using the monitor spectral readings and
the Stockman-Sharpe 2 degree cone fundamentals (Stockman \& Sharpe,
2000). The maximum displayable cone contrasts on our system were 0.1 for
L-M, and 0.88 for S-(L+M). The stimuli in Experiment 3 were temporal
modulations of luminance applied to a disc made using the same raised
cosine envelope as described above, but with no further spatial
modulation. The stimuli counterphase flickered sinusoidally at 4Hz (see
examples in the lower portion of Figure~\ref{fig-examplestim}). In all
experiments, we displayed a binocular fusion lock, consisting of three
concentric rings of small square elements with random colour. A black
central fixation cross was also displayed throughout.

@@

\centering{

\includegraphics[width=0.5\textwidth,height=\textheight]{Figures/examplestim.pdf}

@@

\end{figure}%

@@

\subsection{Procedure}\label{procedure}

@@

\centering{

\includegraphics{Figures/isosettings.pdf}

@@

\caption{\label{fig-isofig}Isoluminance settings from all participants
in Experiments 1 and 2. Panel (a) shows L-M and panel (b) shows S-(L+M)
settings that were subsequently used to generate stimuli in the main
experiments. Within each panel, solid lines show the mean settings for
each participant, and black curves show the range of possible stimuli
displayed during the adjustment task.}

\end{figure}%

@@

\subsection{Data analysis and computational
modelling}\label{data-analysis-and-computational-modelling}

@@

\begin{equation}\phantomsection\label{eq-s1l}{\mathrm{Stage1}_L = \frac{C_L^m}{S + C_L + \omega C_R},}\end{equation}

\begin{equation}\phantomsection\label{eq-s1r}{\mathrm{Stage1}_R = \frac{C_R^m}{S + C_R + \omega C_L},}\end{equation}

\begin{equation}\phantomsection\label{eq-binsum}{\mathrm{binsum} = \mathrm{Stage1}_L + \mathrm{Stage1}_R,}\end{equation}

\begin{equation}\phantomsection\label{eq-s2}{\mathrm{Stage2} = \frac{\mathrm{binsum}^p}{Z + \mathrm{binsum}^q},}\end{equation}
